First of all, none of this would have been possible without Tony Belpaeme. Thank you Tony for giving me the opportunity to start this journey and meet so many people, for introducing me to research, for your enthusiasm, and for making me leave each meeting more motivated than when starting. Thank you also Paul Baxter, for your quiet but sharp comments and ideas. % and for your sense of self-deprecation,
To S\'{e}verin Lemaignan for your energy and positivity and your craving for technical complexity. Thank you to the three of you, for guiding me through these 4 years.

Special thanks to James Kennedy for showing me what a PhD journey could be, for these 2 years and half sharing the office, and for an `interesting' week in Romania to make DREAM work. Thank you to the DREAM team, for the fun of sharing meetings and the opportunity to collaborate on a bigger project, and especially to Hoang-Long Cao and Pablo Gomez for our weekly meetings. To Carlos Cifuentes, Marcela Munera and Jonathan Casas for a warm welcome in Colombia and a productive collaboration. To Madeleine Bartlett for long weeks in schools, repeating the same sentences to more than 100 children and teaching the robot. Also thanks to Charlotte Edmunds, Madeleine Bartlett, Chris Wallbridge, Daniel Hernandez, and Thomas Colin for accepting to proofread the whole thing, dealing with the frenchness of my spelling and grammar, making it definitely better! To everybody not mentioned yet and who has been part of the HRI Plymouth team: Robin Read, Bahar Irfan, Fotios Popadopoulos, and Serge Thill; to the Cognovians for the warm welcome on my arrival in Plymouth and people in CRNS for being a fun and vibrant research group.

The last thoughts are for the ones who gave me the energy to enjoy working on this thesis: people in North Road East who made the house a pleasant and welcoming place I was always happy to return in the evenings, and friends in Plymouth and around the world (Pierre-Henri, Thomas and Tunvez to cite only a few) who provided me with pleasant distractions.

Final thanks to my family and Yan for supporting me through the highs and lows and making me feel appreciated at all times.